---
title: "Sequence Similarity Screen"
editor: visual
author: "Joe Boktor"
date: '2023-01-24'
format: 
  html:
    font-family: helvetica neueu
    page-layout: full
    toc: true
    toc-location: left
    toc-depth: 3
    self-contained: true
    code-fold: true
    code-tools: true
    fig-align: center
---

# Introduction & Background

Engineered immune cell therapies to selectively modulate lymphocytes or myeloid cells of interest are a promising approach for the treatment of cancer, autoimmune diseases, and other disorders (REFS). A key role of the human immune system is to protect the body from microbial pathogens and parasites (REFS). While the human immune system has evolved to recognize and respond to microbial components and pathogens (REFS), microbial pathogens have simultaneously evolved to evade immune recognition and repress inflammatory cell signaling cascades (REFS). Within the last \_\_ years, \_\_\_ additional microbial genomes have been sequenced due to an increase in accessibility to next generate sequencing (NGS) (REFS). Thus, microbial and parasitic interactions with the human immune system may provide an interesting opportunity to develop novel immune therapeutics.

The human body is host to a diverse community of microbes and parasites. It is estimated that \_\_ microbes colonize the human body. Within the gastrointestinal tract, the human body will pass up to \_\_ grams of microbial mass. While the innate immune systems is present in \_ branch of the phylogenetic tree\_\_ the adaptive immune system is detectable within the \_\_\_**branch**\_\_\_ thought to have evolved over \_\_ years ago.

Our hypothesis is that over \_\_ years, microbes and the human immune system have evolved such that \_\_\_. In this project we will explore the hypothesis that human associated microbes and parasites have evolved mechanisms modulate the human immune system. our analysis will mine microbial and parasitic genomic databases to search for 1) signatures of mimetic microbial proteins which resemble immune proteins of interest, and 2) microbial proteins which may be targets for interaction with immune proteins which may target inhibitory/excititory pathway components.

Inspiration for this project arises from two instances. The first of which is the discovery of a Transforming Growth Factor Beta (TGF-β) analog within parasitic roundworms (REF). In fact, there are a number of suspected parasitic homologs to human immune proteins that are known to modulate the immune system.

The second is the discovery of microbial proteins which interfere with proper ITIM and ITAM signalling (REF). This findings suggests that microbial proteins may interfere with the proper functioning of the immune system.

The discover of microbial proteins capable of modulating the immune system provides an exciting opportunity to develop novel therapeutics.

## Objectives

The goal of our project will be to identify microbial proteins with any of the following features 1. Molecular mimicry of human immune proteins involved in immune cell signaling 2. Novel protein interactions with host signaling proteins that can modulate inhibitory or excititory immune signaling pathways.

The goal of this notebook will be to determine which host immune cell proteins are most likely targets for molecular mimicry by microbial proteins, and/or targets for interaction by microbial proteins. Proteins selected in this notebook will be used to search microbial and parasitic genomic databases for homologs or paralogs.

microbial proteins may be targets for interaction with immune proteins.

At a high-level, we are searching for protein that may be targets for interaction with microbial components.

Proteins of interest will likely display some of the following traits:

-   Secretion, Surface Receptors

-   Modulation in immune cells upon activation

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
